---
header-includes:
- \usepackage{xeCJK} 
- \usepackage{fontspec} 
- \setCJKmainfont{微軟正黑體} 
- \XeTeXlinebreaklocale "zh"
- \XeTeXlinebreakskip = 0pt plus 1pt
title: "Statistical Computing HW1"
author: "106033233 資工21 周聖諺"
date: "3/21/2021"
output:
  pdf_document: 
   latex_engine: xelatex
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1:

### (a) Generate standard normal distribution by using Box-Muller approach with 10000 samples. Display the result by the histogram and the boxplot.

**Pseudo Code:**

Step 1. Generate $U_1$, $U_2$ from uniform $U(0, 1)$ independently

Step 2. Let variable 

$$X = \sqrt{-2 ln \ U_1} \cos(2 \pi U_2)$$
$$Y = \sqrt{-2 ln \ U_1} \sin(2 \pi U_2)$$

Step 3. Return $X$ or $Y$, since $X, Y \stackrel{i.i.d}{\sim} N(0,1)$

```{r, echo=TRUE, results='hide'}
library(compositions)

normal_box_muller <- function(n){
  res <- vector("numeric", length=n)
  
  for (i in 0:n) {
    u1 <- runif(1, 0, 1)
    u2 <- runif(1, 0, 1)
    
    radius <- sqrt(-2 * log(u1))
    angle <- 2 * pi * u2
    
    x <- radius * cos(angle)
    y <- radius * sin(angle)
    
    #print(x)
    #print(y)
    
    res[i] <- x
  }
  
  return(res)
}
n <- 10000
res <- normal_box_muller(n)

hist(res, main="Standard Normal with Box-Muller Approach", xlab="Values", breaks=50, freq = FALSE)
boxplot(res, main="Standard Normal with Box-Muller Approach", ylab="Values", freq = FALSE)

r_dist <- rnorm(n, 0, 1)
hist(r_dist, main="Standard Normal with rnorm()", xlab="Values", breaks=50, freq = FALSE)
boxplot(r_dist, main="Standard Normal with rnorm()", ylab="Values", freq = FALSE)
```

### (b) Generate standard normal distribution by using Acceptance and Rejection approach with 10000 samples. Display the result by the histogram and the boxplot.

**Pseudo Code Of Generating Exponential Distribution**

For $X \sim Exp(\lambda)$

Step 1. Generate $U \sim U(0, 1)$

Step 2. Return $-\frac{1}{\lambda} log U$

**Pseudo Code Of Generating Normal Distribution with Acceptance-Rejection Method:**

For $X \sim N(0, 1)$

Step 1. Generate $Y \sim Exp(1)$, $U_1 \sim U(0,1)$

Step 2. If $U_1 \leq \frac{f_{|X|}(Y)}{cg(X)} = e^{-(Y-1)^2}$, set $X = Y$. Otherwise, go back to Step 1.

Step 3. Generate $U_2 \sim U(0,1)$. If $U_2 \leq 0.5$, set $X = |X|$. Otherwise, $X = -|X|$.

Step 4. Return X

```{r, echo=TRUE}
exponential <- function(n, lambda){
  res <- vector("numeric", length=n)

  for (i in 1:n) {
    u <- runif(1, 0, 1)
    
    res[i] <- -(1/lambda) * log(u)
  }
  
  return(res)
}

normal_acc_rej <- function(n){
  res <- vector("numeric", length=n)
  total_num <- 0
  acc_num <- 0

  for (i in 1:n) {
    y <- exponential(1, 1)
    u1 <- runif(1, 0, 1)
    u2 <- runif(1, 0, 1)
    x <- 0
    total_num <- total_num + 1
    
    while (!(u1 <= exp(-((y - 1)**2) / 2))) {
      y <- exponential(1, 1)
      u1 <- runif(1, 0, 1)
      u2 <- runif(1, 0, 1)
      total_num <- total_num + 1
    }
    # Accept
    x <- y
    acc_num <- acc_num + 1
    
    if(u2 <= 0.5){
      x = abs(x)
    }else{
      x = -abs(x)
    }
    
    res[i] <- x
  }
  
  print("Acceptance Rate(%)")
  print(100*acc_num/total_num)
  
  return(res)
}

#n <- 10000
res <- normal_acc_rej(n)

hist(res, main="Standard Normal with Accept-Rejection Approach", xlab="Values", breaks=50, freq = FALSE)
boxplot(res, main="Standard Normal with Accept-Rejection Approach", ylab="Values", freq = FALSE)

r_dist <- rnorm(n, 0, 1)
hist(r_dist, main="Standard Normal with rnorm()", xlab="Values", breaks=50, freq = FALSE)
boxplot(r_dist, main="Standard Normal with rnorm()", ylab="Values", freq = FALSE)
```

## Problem 2:

### (a) Generate Poisson distribution with 10000 samples. Display the result by the histogram and the boxplot.

$$
X \sim Poisson(\mu = 10)
$$

where $\lambda$ the happening rate of the event during $T$ time and the $\mu$ means the average occurrence of the event during $T$ time.

$$
\lambda \cdot T = \mu
$$

**Pseudo Code**

For $Poisson(\mu)$

Step 1. Let $t = 0, X = 0$

Step 2. If $t \leq \mu$, generate $U \sim U(0,1)$. Otherwise, go to Step 5.

Step 3. $t = t - log(U)$

Step 4. if $t \leq \mu$, $X = X + 1$. Otherwise, go back to Step 2.

Step 5. Return $X$

```{r, echo=TRUE}
poisson <- function(n, mu){
  res <- vector("numeric", length=n)

  for (i in 1:n) {
    T <- mu
    
    t <- 0
    x <- 0
    
    while (t <= T) {
      u <- runif(1, 0, 1)
      # lambda = 1
      t <- t - log(u)
      
      if(t <= as.numeric(T)){
        x <- x + 1
      }
    }
    
    res[i] <- x
  }
  
  return(res)
}

#n <- 10000
mu <- 10
res <- poisson(n, mu)

hist(res, main="Poisson Distribution Manual", xlab="Values", freq = FALSE, breaks=25)
boxplot(res, main="Poisson Distribution Manual", ylab="Values", freq = FALSE)

r_dist <- rpois(n, mu)
hist(r_dist, main="Poisson Distribution with rpois()", xlab="Values", breaks=25, freq = FALSE)
boxplot(r_dist, main="Poisson Distribution with rpois()", ylab="Values", freq = FALSE)
```

### (b) Generate Gamma distribution with 10000 samples. Display the result by the histogram and the boxplot.

$X \sim Gamma(\alpha = 5, \beta = 3)$

**Pseudo Code**

For $Gamma(\alpha, \beta)$

Step 1. Generate $X_1, X_2, ..., X_{\alpha} \stackrel{i.i.d}{\sim} Exp(\beta)$

Step 2. Return $\sum_{i = 1}^{\alpha} X_i$

```{r, echo=TRUE}
gamma <- function(n, alpha, beta){
  res <- vector("numeric", length=n)

  for (i in 0:n) {
    # u <- runif(alpha, 0, 1)
    # y <- vector("numeric", length=alpha)
    
    # for (i in 0:alpha) {
      #y[i] <- -1 / beta * log(u[i])
    #}
    #res[i] <- sum(y)
    
    res[i] = sum(exponential(alpha, beta))
  }
  
  return(res)
}

#n <- 10000
alpha <- 5
beta <- 3
res <- gamma(n, alpha, beta)

hist(res, main="Gamma Distribution Manual", xlab="Values", freq = FALSE, breaks=50)
boxplot(res, main="Gamma Distribution Manual", ylab="Values", freq = FALSE)

r_dist <- rgamma(n, shape=alpha, rate=beta)
hist(r_dist, main="Gamma Distribution with rgamma()", xlab="Values", breaks=50, freq = FALSE)
boxplot(r_dist, main="Gamma Distribution with rgamma()", ylab="Values", freq = FALSE)
```

## Problem 3

### (a)

Suppose

$$
X | \mu \sim Poisson(\mu) = \frac{\mu^x e^{-\mu}}{x!}
$$

$$
\mu \sim Gamma(\alpha, \beta) = \frac{\mu^{\alpha - 1} e^{-\frac{\mu}{\beta}}}{\beta^{\alpha} \Gamma(\alpha)}
$$

The marginal distribution $f_X(x)$ of $X$ is 

$$
f_X(x) = \int_{\mu} p(X, \mu) \ d\mu = \int_{\mu} p(X|\mu)p(\mu) \ d\mu
$$

$$
= \int_{\mu} \frac{\mu^x e^{-\mu}}{x!} \cdot \frac{\mu^{\alpha - 1} e^{-\frac{\mu}{\beta}}}{\beta^{\alpha} \Gamma(\alpha)} \ d\mu
$$

$$
= \frac{1}{x! \Gamma(\alpha) \beta^{\alpha}} \int_0^{\infty} \mu^x e^{-\mu} \mu^{\alpha - 1} e^{-\frac{\mu}{\beta}} \ d\mu
$$

$$
= \frac{1}{x! \Gamma(\alpha) \beta^{\alpha}} \int_0^{\infty} \mu^{\alpha + x - 1} e^{-\mu (1 + \frac{1}{\beta})} \ d\mu
$$

$$
= \frac{1}{\Gamma(x + 1) \Gamma(\alpha) \beta^{\alpha}} \Gamma(\alpha + x) \frac{\beta}{1 + \beta}
$$

$$
= \begin{pmatrix}
\alpha - 1 + x\\
x
\end{pmatrix}
(\frac{1}{1+\beta})^{\alpha} \ 
(1 - \frac{1}{1+\beta})^{x}
$$

Let $n = \alpha, p = \frac{1}{1 + \beta}$

$$
= \begin{pmatrix}
n - 1 + x\\
x
\end{pmatrix} p^n (1 - p)^{x}
$$

It is a Negative Binomial distribution $\mathcal{NB}(n, p)$

**Pseudo Code Of Geometric **

For $Geo(p)$

Step 1. Generate $U \sim U(0, 1)$

Step Return $\lfloor \frac{log \ U}{log \ (1 - p)} \rfloor$

**Pseudo Code Of Negative Binomial**

For $NB(n, p)$

Step 1. Generate $X_1, X_2, ..., X_{n} \stackrel{i.i.d}{\sim} Geo(p)$

Step 2. Return $\sum_{i = 1}^{n} X_i$

### (b)

```{r, echo=TRUE}
geo <- function(n, p){
  res <- vector("numeric", length=n)

  for (i in 1:n) {
    u <- runif(1, 0, 1)
    #print(u)
    #print(log(u))
    #print(log(1 - p))
    res[i] <- floor(log(u) / log(1 - p))
  }
  
  return(res)
}

nb <- function(n, m, p){
  res <- vector("numeric", length=n)

  for (i in 1:n) {
    geo_res <- vector("numeric", length=m)
    
    geo_res <- geo(m, p)
    
    #print(geo_res)
    
    res[i] <- sum(geo_res)
  }
  
  return(res)
}

#n <- 10000
alpha <- 5
beta <- 3
res <- nb(n, alpha, 1/(1 + beta))

#print(res)

hist(res, main="Negative Binomial Distribution Manual", xlab="Values", freq = FALSE, breaks=50)
boxplot(res, main="Negative Binomial Distribution Manual", ylab="Values", freq = FALSE)

r_dist <- rnbinom(n, alpha, 1/(1 + beta))
hist(r_dist, main="Negative Binomial Distribution with rnbinom()", xlab="Values", breaks=50, freq = FALSE)
boxplot(r_dist, main="Negative Binomial Distribution with rnbinom()", ylab="Values", freq = FALSE)
```

### (c) What are the mean and variance of $X$?

Mean 

$$
\frac{pr}{1-p} = \frac{\frac{1}{1 + \beta} \alpha}{1 - \frac{1}{1+\beta}} = \frac{\alpha}{\beta} = \frac{5}{3}
$$

Variance

$$
\frac{pr}{(1-p)^2} = \frac{\frac{1}{1 + \beta} \alpha}{(1 - \frac{1}{1+\beta})^2} = \frac{\alpha (1 + \beta)}{\beta^2} = \frac{20}{9}
$$

# Problem 4

### (a)

$$
X_1 \sim \mathcal{N}(\mu_1, \sigma_1^2), \ X_2 \sim \mathcal{N}(\mu_2, \sigma_2^2)
$$

A mixture model of $X_1, X_2$

$$
f_{X_1, X_2}(x) = p_1 \cdot p_{X_1}(x) + p_2 \cdot p_{X_2}(x)
$$

$$
= p_1 \cdot \frac{1}{\sigma_1 \sqrt{2 \pi}} e^{- \frac{1}{2} (\frac{x-\mu_1}{\sigma_1})^2} + p_2 \cdot \frac{1}{\sigma_2 \sqrt{2 \pi}} e^{- \frac{1}{2} (\frac{x-\mu_2}{\sigma_2})^2}
$$

Let $\mu_1 = 0, \mu_2 = 3$ and $\sigma_1^2 = \sigma_2^2 = 1$

$$
= p_1 \cdot \frac{1}{\sqrt{2 \pi}} e^{- \frac{1}{2} x^2} + (1 - p_1) \cdot \frac{1}{\sqrt{2 \pi}} e^{- \frac{1}{2} (x-3)^2}
$$

### (b)

Let $p_1 = 0.75$ and generate 10000 samples from the mixture model.

```{r, echo=TRUE}
mix_acc_rej <- function(n, p_1, mu_1, mu_2, sigma_1, sigma_2){
  res <- vector("numeric", length=n)

  for (i in 0:n) {
    p <- runif(1, 0, 1)
    shift <- 0
    scale <- 0
    if(p <= p_1){
      shift <- mu_1
      scale <- sigma_1
    }else{
      shift <- mu_2
      scale <- sigma_2
    }
    
    y <- exponential(1, 1)
    u1 <- runif(1, 0, 1)
    u2 <- runif(1, 0, 1)
    x <- 0
    
    while (!(u1 <= exp(-((y - 1)**2) / 2))) {
      y <- rexp(1, 1)
      u1 <- runif(1, 0, 1)
      u2 <- runif(1, 0, 1)
    }
    # Accept
    x <- y
    
    if(u2 <= 0.5){
      x = abs(x)
    }else{
      x = -abs(x)
    }
    
    x <- x * scale + shift
    
    res[i] <- x
  }
  
  return(res)
}

#n <- 10000
res <- mix_acc_rej(n, 0.75, 0, 3, 1, 1)

hist(res, main="Mixed Gaussian with Accept-Rejection Approach", xlab="Values", breaks=100, freq = FALSE)
```

The distribution seems bimodal.